\Title

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
perspectives

%Note that this supervised learning scheme can be extended to a variety of tasks. It would follow from the emergence of new kernels adapted to this new task after supervised learning.  This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules (see for instance~\citep{benosman_asynchronous_2012} for computing the optic flow using the luminance conservation rule). We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks from motion extraction, optic flow or time-to-contact maps.
% >>>>>>>>>>> LuP is here <<<<<<<<<<<

\backmatter

>>>>

%Note that this supervised learning scheme can be extended to a variety of tasks. It would follow from the emergence of new kernels adapted to this new task after supervised learning.  This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules (see for instance~\citep{benosman_asynchronous_2012} for computing the optic flow using the luminance conservation rule). We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks from motion extraction, optic flow or time-to-contact maps.
% >>>>>>>>>>> LuP is here <<<<<<<<<<<
In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image. This gradient is defined in both horizontal and vertical dimensions and its maxima are generally oriented. Taken together, these oriented edges form the contours of visual objects in the scene (Koenderink). Thus, there is a dependence within the event stream between motion and orientation information. It would be crucial to study this interdependence further, which could be initiated by training the model on a dataset with orientation labels. Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them. In particular it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem (Perrinet 2012).

The model is trained on a low-level local motion detection task, and one might wonder if it could be trained on higher-level tasks. An example task would be depth estimation in the visual scene. There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue (Rogers). This is because objects close to an observer move relatively faster on the retina than an object at a great distance. Using this information, one can segment objects and estimate their depth (Yoonessi). However, this would first require computing the optical flow. A possible implementation is therefore to add a new layer to our model, in analogy with the hierarchical organization highlighted in the visual cortex. This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension as an output label. This could have direct and important applications, for example in autonomous driving, to detect obstacles in a fast and robust way. Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates notably to resolve uncertain estimates (Nawrot).

In conclusion, the model we have presented provides a way to efficiently process event-based signals. We have shown that we can train the model in a semi-supervised manner, knowing the output labels but not knowing when they occur. Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data (Barlow). This type of learning is thought to be predominant in the central nervous system and, assuming that the signal is sparse (Olshausen), one could extend those hebbian sparse learning schemes to spikes (Perrinet Masquelier). This would be particularly well suited to the exploration of neurobiological data. In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains (Ikegaya), packets (Luczak), or hippocampal sequences (Pastalkora 2008 Villette 2015). These patterns are stereotypical and robust, as they can be activated in the same motif from day to day (Haimerl 2019). In contrast to conventional methodologies that are used to process neurobiological data, such a event-based model would be able to answer key questions regarding the representation of information in neurobiological data. Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.

%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter
